---
title: "D3 Treemap"
author: "Ming Chen"
date: "7/5/2017"
output: html_document
---

<script src="https://d3js.org/d3.v4.min.js"></script>
<svg width="960" height="570"></svg>

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
library(dplyr)
```




```{css}
form {
  font-family: "Helvetica Neue", Helvetica, Arial, sans-serif;
}

svg {
  font: 10px sans-serif;
}
```


```{r echo=FALSE}
# add radio buttons
input_1 = tags$input('Size') %>% 
  tagAppendAttributes(
    type="radio",
    name="mode",
    value="sumBySize",
    checked=TRUE
  )
input_2 = tags$input('Count') %>% 
  tagAppendAttributes(
    type="radio",
    name="mode",
    value="sumByCount"
  )
tags$form(
  list(tags$label(input_1), tags$label(input_2))
)
```


A treemap recursively subdivides area into rectangles; the area of any node in the tree corresponds to its value. This example uses color to encode different packages of the Flare visualization toolkit. Treemap design invented by [Ben Shneiderman](http://www.cs.umd.edu/~ben/). Squarified algorithm by [Bruls, Huizing and van Wijk](http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.36.6685). Data courtesy [Jeff Heer](http://flare.prefuse.org/).


```{r echo=FALSE}
# add d3 min js
d3_min_js = tags$script() %>%
  tagAppendAttributes(src='https://d3js.org/d3.v4.min.js')
```



```{js}
var svg = d3.select("svg"),
    width = +svg.attr("width"),
    height = +svg.attr("height");

var fader = function(color) { return d3.interpolateRgb(color, "#fff")(0.2); },
    color = d3.scaleOrdinal(d3.schemeCategory20.map(fader)),
    format = d3.format(",d");

var treemap = d3.treemap()
    .tile(d3.treemapResquarify)
    .size([width, height])
    .round(true)
    .paddingInner(1);

d3.json("data/flare.json", function(error, data) {
  if (error) throw error;

  var root = d3.hierarchy(data)
      .eachBefore(function(d) { d.data.id = (d.parent ? d.parent.data.id + "." : "") + d.data.name; })
      .sum(sumBySize)
      .sort(function(a, b) { return b.height - a.height || b.value - a.value; });

  treemap(root);

  var cell = svg.selectAll("g")
    .data(root.leaves())
    .enter().append("g")
      .attr("transform", function(d) { return "translate(" + d.x0 + "," + d.y0 + ")"; });

  cell.append("rect")
      .attr("id", function(d) { return d.data.id; })
      .attr("width", function(d) { return d.x1 - d.x0; })
      .attr("height", function(d) { return d.y1 - d.y0; })
      .attr("fill", function(d) { return color(d.parent.data.id); });

  cell.append("clipPath")
      .attr("id", function(d) { return "clip-" + d.data.id; })
    .append("use")
      .attr("xlink:href", function(d) { return "#" + d.data.id; });

  cell.append("text")
      .attr("clip-path", function(d) { return "url(#clip-" + d.data.id + ")"; })
    .selectAll("tspan")
      .data(function(d) { return d.data.name.split(/(?=[A-Z][^A-Z])/g); })
    .enter().append("tspan")
      .attr("x", 4)
      .attr("y", function(d, i) { return 13 + i * 10; })
      .text(function(d) { return d; });

  cell.append("title")
      .text(function(d) { return d.data.id + "\n" + format(d.value); });

  d3.selectAll("input")
      .data([sumBySize, sumByCount], function(d) { return d ? d.name : this.value; })
      .on("change", changed);

  var timeout = d3.timeout(function() {
    d3.select("input[value=\"sumByCount\"]")
        .property("checked", true)
        .dispatch("change");
  }, 2000);

  function changed(sum) {
    timeout.stop();

    treemap(root.sum(sum));

    cell.transition()
        .duration(750)
        .attr("transform", function(d) { return "translate(" + d.x0 + "," + d.y0 + ")"; })
      .select("rect")
        .attr("width", function(d) { return d.x1 - d.x0; })
        .attr("height", function(d) { return d.y1 - d.y0; });
  }
});

function sumByCount(d) {
  return d.children ? 0 : 1;
}

function sumBySize(d) {
  return d.size;
}
```


---

# Treemap

This is an example of generating the same [Treemap](https://bl.ocks.org/mbostock/4063582) in R markdown by integrating `CSS`, `Javascript`, `R script` and `R markdown` into a single R markdown document.

# Render CSS

```
form {
  font-family: "Helvetica Neue", Helvetica, Arial, sans-serif;
}

svg {
  font: 10px sans-serif;
}
```

# Render radio buttons

```{r eval=FALSE}
# add radio buttons
input_1 = tags$input('Size') %>% 
  tagAppendAttributes(
    type="radio",
    name="mode",
    value="sumBySize",
    checked=TRUE
  )
input_2 = tags$input('Count') %>% 
  tagAppendAttributes(
    type="radio",
    name="mode",
    value="sumByCount"
  )
tags$form(
  list(tags$label(input_1), tags$label(input_2))
)
```

# Add text

```
A treemap recursively subdivides area into rectangles; the area of any node in the tree corresponds to its value. This example uses color to encode different packages of the Flare visualization toolkit. Treemap design invented by [Ben Shneiderman](http://www.cs.umd.edu/~ben/). Squarified algorithm by [Bruls, Huizing and van Wijk](http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.36.6685). Data courtesy [Jeff Heer](http://flare.prefuse.org/).
```

# Attach d3.min.js

```{r eval=FALSE}
# add d3 min js
d3_min_js = tags$script() %>%
  tagAppendAttributes(src='https://d3js.org/d3.v4.min.js')
```

# Add plotting Javascript

```
var svg = d3.select("svg"),
    width = +svg.attr("width"),
    height = +svg.attr("height");

var fader = function(color) { return d3.interpolateRgb(color, "#fff")(0.2); },
    color = d3.scaleOrdinal(d3.schemeCategory20.map(fader)),
    format = d3.format(",d");

var treemap = d3.treemap()
    .tile(d3.treemapResquarify)
    .size([width, height])
    .round(true)
    .paddingInner(1);

d3.json("data/flare.json", function(error, data) {
  if (error) throw error;

  var root = d3.hierarchy(data)
      .eachBefore(function(d) { d.data.id = (d.parent ? d.parent.data.id + "." : "") + d.data.name; })
      .sum(sumBySize)
      .sort(function(a, b) { return b.height - a.height || b.value - a.value; });

  treemap(root);

  var cell = svg.selectAll("g")
    .data(root.leaves())
    .enter().append("g")
      .attr("transform", function(d) { return "translate(" + d.x0 + "," + d.y0 + ")"; });

  cell.append("rect")
      .attr("id", function(d) { return d.data.id; })
      .attr("width", function(d) { return d.x1 - d.x0; })
      .attr("height", function(d) { return d.y1 - d.y0; })
      .attr("fill", function(d) { return color(d.parent.data.id); });

  cell.append("clipPath")
      .attr("id", function(d) { return "clip-" + d.data.id; })
    .append("use")
      .attr("xlink:href", function(d) { return "#" + d.data.id; });

  cell.append("text")
      .attr("clip-path", function(d) { return "url(#clip-" + d.data.id + ")"; })
    .selectAll("tspan")
      .data(function(d) { return d.data.name.split(/(?=[A-Z][^A-Z])/g); })
    .enter().append("tspan")
      .attr("x", 4)
      .attr("y", function(d, i) { return 13 + i * 10; })
      .text(function(d) { return d; });

  cell.append("title")
      .text(function(d) { return d.data.id + "\n" + format(d.value); });

  d3.selectAll("input")
      .data([sumBySize, sumByCount], function(d) { return d ? d.name : this.value; })
      .on("change", changed);

  var timeout = d3.timeout(function() {
    d3.select("input[value=\"sumByCount\"]")
        .property("checked", true)
        .dispatch("change");
  }, 2000);

  function changed(sum) {
    timeout.stop();

    treemap(root.sum(sum));

    cell.transition()
        .duration(750)
        .attr("transform", function(d) { return "translate(" + d.x0 + "," + d.y0 + ")"; })
      .select("rect")
        .attr("width", function(d) { return d.x1 - d.x0; })
        .attr("height", function(d) { return d.y1 - d.y0; });
  }
});

function sumByCount(d) {
  return d.children ? 0 : 1;
}

function sumBySize(d) {
  return d.size;
}
```

