---
title: "Geom Layer"
author: "Ming Chen"
date: "6/25/2017"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
library(ggplot2)
```


# Geom Layer

## `geom_abline`, `geom_hline` and `geom_vline`

```{r}
ggplot(data = mtcars) +
  geom_point(mapping = aes(x=0, y=5), 
             stat = "identity", 
             position = "identity", 
             col = "red") +
  geom_abline(intercept = 5, slope = 2) +
  geom_hline(yintercept = 2.5, col = "blue") +
  geom_vline(xintercept = 5.0, col = "red") +
  geom_text(mapping = aes(x=2, y=5), 
            stat = "identity",
            position = "identity",
            label = "intercept = 5, slop = 2") +
  geom_text(mapping = aes(x=2, y=1.5),
            stat = "identity",
            position = "identity",
            label = "yintercept = 2.5") + 
  geom_text(mapping = aes(x=5.2, y=7.2),
            stat = "identity",
            position = "identity",
            label = "xintercept = 5.0",
            angle = 270) +
  coord_cartesian(xlim = c(0,10), ylim = c(0, 10)) + 
  facet_grid(am~.)

```

## `geom_bar` and `geom_col`

* `geom_bar` uses `stat_count` to calculate the y values by default.
* `geom_col` uses `stat_identity` by default. You will need to provide the `y` mapping.

```{r}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x=class, fill=factor(cyl)),
           position = position_stack()) +
  coord_flip()

my_data = data.frame(x = rep(LETTERS[1:5], each=3),
                     y = sample(1:100, 15),
                     class = rep(letters[1:3], 5))
ggplot(data = my_data) +
  geom_col(mapping = aes(x=x, y=y, fill=class),
           position = position_stack()) +
  coord_flip()
```

## `geom_bind2d`

* `geom_bind2d` maps x, y and z to a coordinate system. z is the counts for each xy combination. If x and y are numeric, x and y are split into series of intervals (**bin**).

```{r}
ggplot(data = mpg) +
  geom_bin2d(mapping = aes(x=manufacturer, y=displ)) +
  theme(axis.text.x = element_text(angle = 90))
```

