---
title: "ggplot2 Grammar"
author: "Ming Chen"
date: "6/25/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
library(ggplot2)
```

## The grammar of ggplot2 plots

The following ggplot template comes from the book ***[R for Data Science](http://r4ds.had.co.nz/)*** by Garrett Grolemund & Hadley Wickham.

```
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

Each ggplot2 plot is a combination use of seven parameters as you can in the template.

* `DATA`: contains values presented on a plot.
* `MAPPINGS`: determines how values are mapped to plot elements.
* `STAT`: the rule of transforming original values to new values and then mapping new values to plot elements. 
* `POSITION`: arranges locations of observations presented on a plot.
* `COORDINATE_FUNCTION`: provides several different coordinate systems.
* `FACET_FUNCTION`: split plot into facets.

## Examples

```{r}
ggplot(data = mtcars) +                           
  geom_point(mapping = aes(x = mpg, y = disp),  
             stat = 'identity',
             position = 'identity') +
  coord_cartesian() + 
  facet_grid(~am)
```


